---
title: "Northern Norways New Companies"
author: "Lucas Stark"
date: "01 10 2019"
output: 
  html_document:
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(plyr) #for ddply()
library(tidyverse)
library(rvest)

```

# Research Question

The following research was conducted to answer the following question: "How many new companies were registered in Northern Norway per county and month in 2019?"

# Data Source
```{r importing, include=FALSE}
#set url link
link <- read_html("https://w2.brreg.no/kunngjoring/kombisok.jsp?datoFra=01.01.2019&datoTil=30.09.2019&id_region=100&id_fylke=-+-+-&id_niva1=2&id_bransje1=0")

#prepare and import
dat_pre <- html_nodes(link, xpath = "//table")
dat <- as.data.frame(html_table(html_nodes(dat_pre, "table")))
```


The data was imported from the website of [Bronnoysundregistrene of Norway](https://w2.brreg.no/kunngjoring/kombisok.jsp?datoFra=01.01.2019&datoTil=30.09.2019&id_region=100&id_fylke=-+-+-&id_niva1=2&id_bransje1=0).

# Data Cleaning

```{r first cleaning, include=FALSE}
#get each data set for every country
#there is some rubbish in the beginning and then a heading over every of the countys entries
#name variables
names(dat) <- c("x1","company","x2","number","x3","date","x4","ann_type","x5")

#get rid of first unneccessary columns
dat <- dat[,c("company","x2","number","date","ann_type")]

#get line numbers that seperate the counties
dat$x2 <- as.factor(dat$x2)
dat$x2[dat$x2==""] <- NA
summary(dat$x2)
x <- which(is.na(dat$x2)==F)
x <- c(x,length(dat$x2)+1) #add line of end of df (plus one for following command)

#split dataframe in frames for each county
dat_list <- as.list(NA)
for(i in 1:(length(x)-1)){
  dat_sub <- dat[(x[i]+1):(x[i+1]-1),]
  dat_sub <- as.data.frame(dat_sub)
  dat_list[[i]] <- dat_sub
}

#assign new column to every data frame with county name
for(i in 1:(length(dat_list))){
  dat_list[[i]]$county <- levels(dat_list[[1]]$x2)[i+1]
}

#create new dataset with all existing data
dat2 <- NA
for(i in 1:(length(dat_list))){
  dat2 <- rbind(dat2,dat_list[[i]])
}

```

The data was cleaned in two major steps. The first one focused on getting the dataset for each seperate county of Northern Norway, getting rid of chunk lines and joining the datasets again with a new variable for the county name. 

```{r second cleaning, include=FALSE}
#delete unneccessary second column and reorder them
dat2 <- dat2[,c("county","company","number","date","ann_type")]

#delete all rows with only missing values
dat2 <- dat2[dat2$company!="",]
dat2 <- dat2[-is.na(dat2$company),]

#make date as date variable and others to factors
dat_clean <- dat2 %>%
  mutate(date = as.Date(date,format("%d.%m.%Y")),
         county = as.factor(county),
         company = as.factor(company),
         number = as.factor(number),
         ann_type = as.factor(ann_type))

#delete all "Rettelse av foretaksnavn" (Correction of company name)
dat_clean <- dat_clean[!dat_clean$ann_type=="Rettelse av foretaksnavn",]


#a final check
summary(dat_clean)
str(dat_clean)

#some companies have more than one entry. ?
dat_clean[dat_clean$company=="NAPOLEON EIENDOM AS",]

```

The second step focused on selecting the necessary variables and transforming them into the appropriate type.

In summary the structure of the final and cleaned dataset looks like this:

```{r summary}
str(dat_clean)
```

# Data Analysis

To find out how many companies have been registered per month and per county the following linegraph was created.

```{r Plot}
# Datensatz vorbereiten
dat_graph <- dat_clean %>%
  mutate(month = lubridate::month(date)) %>%
  mutate(month = paste0("2019","-",month,"-","01")) %>%
  mutate(month = lubridate::ymd(month)) %>%
  ddply(c("county", "month"), summarise, n=as.numeric(length(company)))

# Plot kreieren
dat_graph %>%
  ggplot(aes(x=month,y=n, color=county, group=county)) +
  facet_grid(~county) +
  geom_point() +
  geom_line() +
  scale_y_continuous(limits =  c(0,150),
                     breaks=seq(0, 150, by = 10)) +
  scale_x_date(date_labels = "%b",
               date_breaks = "1 month") +
  ggtitle("Number of newly registered companys per county and month in Northern Norway")  + 
  theme_light() +
  theme(legend.position = "none") + 
  theme(axis.title.x=element_blank(), axis.title.y=element_blank()) +
  geom_hline(aes(yintercept = mean(n)), color="grey", linetype="dashed" ) +
  geom_smooth(method = "lm", se = FALSE)

dat_maxmonth <- dat_graph %>% 
  group_by(month) %>% 
  summarise(max_month = mean(n)) %>%
  as.data.frame()

maxmonth <- max(dat_maxmonth$max_month)
```

As you can see, in general, most new companies were registered in `r dat_clean$county[max(summary(dat_clean$county))][1]`.
The overall mean of new registrations per month is `r round(mean(dat_graph$n),2)` which is depicted as the dashed line in the graphic.
The month with the most new registrations is `r lubridate::month(dat_maxmonth[dat_maxmonth$max_month==maxmonth,1][1], label = T, abbr = FALSE)` with an average of `r maxmonth` new registrations.

